---
title: "Intelligence of Dogs"
author: "Kimberly Cable"
date: "05-21-2022"
output:
  pdf_document: default
bibliography: bibliography.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Final Project - Step 2

### How to import and clean my data

```{r echo = FALSE}
## Load the ggplot2 package
library(ggplot2)
library(dplyr)
library(magrittr)
theme_set(theme_minimal())
```


Load and read each of the datasets:

* **dog_intelligence.csv**

```{r echo = FALSE}
intelligence_df <- read.csv("Final_Project/data/dog_intelligence.csv", 
    header = TRUE, 
    stringsAsFactors = FALSE)
head(intelligence_df)
```

* **AKC Breed Info.csv**

```{r echo = FALSE}
breed_df <- read.csv("Final_Project/data/AKC Breed Info.csv")
head(breed_df)
```

* **Table_4_Heterozygosity_85_breeds.csv**

```{r echo = FALSE}
heterozygosity_4_df <- read.csv("Final_Project/data/Table_4_Heterozygosity_85_breeds.csv")
head(heterozygosity_4_df)
```

* **Table_5_Expected_Heterozygosity_60_breeds.csv**

```{r echo = FALSE}
heterozygosity_5_df <- read.csv("Final_Project/data/Table_5_Expected_Heterozygosity_60_breeds.csv")
colnames(heterozygosity_5_df)[2] <- 'Heterozygosity_x10_4' 

head(heterozygosity_5_df)
```


**Create New Dataframe from the Intelligence data**

```{r echo = FALSE}
combined_df <- intelligence_df

head(combined_df)
```


**Inner Join Breed data to new combined df on key Breed**

```{r echo = FALSE}
combined_df <- combined_df %>% 
    inner_join(breed_df, by = c("Breed" = "Breed"))

head(combined_df)
```

**Inner Join Heterozygosity 4 to new combined df on key Breed = Population**

```{r echo = FALSE}
combined_df <- combined_df %>% 
    inner_join(heterozygosity_4_df, by = c("Breed" = "Population"))

head(combined_df)
```

**Inner Join Heterozygosity 5 to new combined df on key Breed**

```{r echo = FALSE}
combined_df <- combined_df %>% 
    inner_join(heterozygosity_5_df, by = c("Breed" = "Breed"))

head(combined_df)
```

**Convert n/a or na to empty cell**

```{r echo = FALSE}
combined_df[combined_df == 'n/a'] <- ''
combined_df[combined_df == 'na'] <- ''
```


**Convert obey to numeric**

```{r echo = FALSE}
combined_df$obey <- gsub("%", "", as.character(combined_df$obey)) 

combined_df$obey <- as.numeric(combined_df$obey) / 100
```

**Convert height and weight to numeric**

```{r echo = FALSE}
combined_df$height_low_inches <- as.numeric(combined_df$height_low_inches)
combined_df$height_high_inches <- as.numeric(combined_df$height_high_inches)
combined_df$weight_low_lbs <- as.numeric(combined_df$weight_low_lbs)
combined_df$weight_high_lbs <- as.numeric(combined_df$weight_high_lbs)
```

### What does the final data set look like?

```{r echo = FALSE}
head(combined_df)
```

### What information is not self-evident?
* Initially I do not know exactly what Heterozygosity and Heterozygosity (x10-4) are and the difference between the two columns.

### What are different ways you could look at this data?
One could strictly look at the obey percentage without looking at the number of reps a dog can do. You can also just look at the upper and lower reps versus taking the average number of reps a dog can do. Same problem with height and weight if I were to look at if intelligence is strictly by the weight of a breed or how tall a breed is.

### How do you plan to slice and dice the data?

* **Add average weight and height to dataframe**

```{r echo = FALSE}
combined_df$avg.weight = rowMeans(combined_df[,c('weight_low_lbs', 'weight_high_lbs')], na.rm = TRUE)
combined_df$avg.height = rowMeans(combined_df[,c('height_low_inches', 'height_high_inches')], na.rm = TRUE)

head(combined_df)
```

### How could you summarize your data to answer key questions?

```{r echo = FALSE}
summary(combined_df)
```

### What types of plots and tables will help you illustrate the findings to your questions?

```{r echo = FALSE}
ggplot(combined_df, aes(obey)) + 
    labs(title = "Histogram Obey Percentage", 
        x = "Obey Percentage", y = "Frequency") +
    geom_histogram(bins = 10, aes(y = ..density..), color = "black", fill = "black") +
    stat_function(fun = dnorm, 
                  args = list(mean = mean(combined_df$obey, na.rm = TRUE), 
                              sd = sd(combined_df$obey, na.rm = TRUE)),
                              color = 'red', size = 1)
```


```{r echo = FALSE}
ggplot(combined_df, aes(avg.weight)) + 
    labs(title = "Histogram Average Weight", 
        x = "Weight (lbs)", y = "Frequency") +
    geom_histogram(bins = 10, aes(y = ..density..), color = "black", fill = "black") +
    stat_function(fun = dnorm, 
                  args = list(mean = mean(combined_df$avg.weight, na.rm = TRUE), 
                              sd = sd(combined_df$avg.weight, na.rm = TRUE)),
                              color = 'red', size = 1)
```


```{r echo = FALSE}
ggplot(combined_df, aes(avg.height)) + 
    labs(title = "Histogram Average Height", 
        x = "Height (inches)", y = "Frequency") +
    geom_histogram(bins = 10, aes(y = ..density..), color = "black", fill = "black") +
    stat_function(fun = dnorm, 
                  args = list(mean = mean(combined_df$avg.height, na.rm = TRUE), 
                              sd = sd(combined_df$avg.height, na.rm = TRUE)),
                              color = 'red', size = 1)
```


```{r echo = FALSE}
ggplot(combined_df, aes(Heterozygosity)) + 
    labs(title = "Histogram Heterozygosity", 
        x = "Heterozygosity", y = "Frequency") +
    geom_histogram(bins = 10, aes(y = ..density..), color = "black", fill = "black") +
    stat_function(fun = dnorm, 
                  args = list(mean = mean(combined_df$Heterozygosity, na.rm = TRUE), 
                              sd = sd(combined_df$Heterozygosity, na.rm = TRUE)),
                              color = 'red', size = 1)
```


```{r echo = FALSE}
ggplot(combined_df, aes(Heterozygosity_x10_4)) + 
    labs(title = "Histogram Heterozygosity_x10_4", 
        x = "Heterozygosity_x10_4", y = "Frequency") +
    geom_histogram(bins = 10, aes(y = ..density..), color = "black", fill = "black") +
    stat_function(fun = dnorm, 
                  args = list(mean = mean(combined_df$Heterozygosity_x10_4, na.rm = TRUE), 
                              sd = sd(combined_df$Heterozygosity_x10_4, na.rm = TRUE)),
                              color = 'red', size = 1)
```

* **Scatter Plot of obey and avg.weight**

```{r echo = FALSE}
scatter <- ggplot(combined_df, aes(obey, avg.weight))
scatter + geom_point() +
    scale_x_continuous(n.breaks = 10) +
    geom_smooth(method = 'lm', alpha = 0.1) +
                    labs(x = "Obey Percentage", y = "Average Weight") +
    ggtitle('Obey Percentage and Average Weight')
```
* **Scatter Plot of obey and Heterozygosity_x10_4**

```{r echo = FALSE}
scatter <- ggplot(combined_df, aes(obey, Heterozygosity_x10_4))
scatter + geom_point() +
    scale_x_continuous(n.breaks = 10) +
    geom_smooth(method = 'lm', alpha = 0.1) +
                    labs(x = "Obey Percentage", y = "Heterozygosity x10.4") +
    ggtitle('Obey Percentage and Heterozygosity x10.4')
```

* **Correlation between obey percentage and avg.weight**

```{r echo = FALSE}
cor.test(combined_df$obey, combined_df$avg.weight, use = "complete.obs")
```
Since the correlation is 0.19 and the p-value is 0.39 we can say that the correlation between the two variables is not significant.  Also, the intervals cross 0 so as one goes up the other goes up but then it is reversed.

* **Correlation between obey percentage and Heterozygosity_x10_4**

```{r echo = FALSE}
cor.test(combined_df$obey, combined_df$Heterozygosity_x10_4, use = "complete.obs")
```
Since the correlation is 0.09 and the p-value is 0.66 we can say that the correlation between the two variables is not significant.  Also, the intervals cross 0 so as one goes up the other goes up but then it is reversed.

* **Correlation between avg.weight and Heterozygosity_x10_4**

```{r echo = FALSE}
cor.test(combined_df$avg.weight, combined_df$Heterozygosity_x10_4, use = "complete.obs")
```

### Do you plan on incorporating any machine learning techniques to answer your research questions? Explain.

```{r echo = FALSE}
combined_model <- lm(obey ~ avg.weight + avg.height + Heterozygosity_x10_4, data = combined_df)

summary(combined_model)
```
Looking at the Adjusted R-squared of -0.37 and all p-values for the variables are not significant it does not look like any other the variables help with the percentage a dog can obey.

### Questions for future steps.

More research would need to be done to find out if any other data can be linked to a dog's intelligence.
