---
title: "Week 8 - 9"
author: "Kimberly Cable"
date: "05-14-2022"
output:
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Assignment 6

```{r echo = FALSE}
## Load the `data/r4ds/heights.csv` to
heights_df <- read.csv("data/r4ds/heights.csv")
head(heights_df)
```

```{r echo = FALSE}
## Load the ggplot2 library
library(ggplot2)
```

## Fit a linear model using the `age` variable as the predictor and `earn` as the outcome

lm(outcome ~ predictors, data = dataframe)
```{r}
age_lm <-  lm(earn ~ age, data = heights_df)
```

## View the summary of your model using `summary()`
```{r}
summary(age_lm)
```

## Creating predictions using `predict()`
```{r}
age_predict_df <- data.frame(earn = predict(age_lm, heights_df), age = heights_df$age)
head(age_predict_df)
```

## Plot the predictions against the original data
```{r}
ggplot(data = heights_df, aes(y = earn, x = age)) +
  geom_point(color = 'blue') +
  geom_line(color='red',data = age_predict_df, aes(y = earn, x = age))
```

```{r}
mean_earn <- mean(heights_df$earn)
mean_earn
```

## Corrected Sum of Squares Total
```{r}
sst <- sum((mean_earn - heights_df$earn)^2)
sst
```

## Corrected Sum of Squares for Model
```{r}
ssm <- sum((mean_earn - age_predict_df$earn)^2)
ssm
```

## Residuals
```{r}
residuals <- heights_df$earn - age_predict_df$earn
head(residuals)
```

## Sum of Squares for Error
```{r}
sse <- sum(residuals^2)
sse
```

## R Squared R^2 = SSM/SST
```{r}
r_squared <- ssm / sst
r_squared
```

## Number of observations
```{r}
n <- nrow(heights_df)
n
```

## Number of regression parameters
```{r}
p <- 2
```

## Corrected Degrees of Freedom for Model (p-1)
```{r}
dfm <- p - 1
```

## Degrees of Freedom for Error (n-p)
```{r}
dfe <- n - p
```

## Corrected Degrees of Freedom Total:   DFT = n - 1
```{r}
dft <- n - 1
```

## Mean of Squares for Model:   MSM = SSM / DFM
```{r}
msm <- ssm / dfm
msm
```

## Mean of Squares for Error:   MSE = SSE / DFE
```{r}
mse <- sse / dfe
mse
```

## Mean of Squares Total:   MST = SST / DFT
```{r}
mst <- sst / dft
mst
```

## F Statistic F = MSM/MSE
```{r}
f_score <- msm / mse
f_score
```

## Adjusted R Squared R2 = 1 - (1 - R2)(n - 1) / (n - p)
```{r}
adjusted_r_squared <- 1 - ((1 - r_squared) * dft) / dfe
adjusted_r_squared
```

## Calculate the p-value from the F distribution
```{r}
p_value <- pf(f_score, dfm, dft, lower.tail = F)
p_value
```
# Assignment 7

## Load the `data/r4ds/heights.csv` to
```{r echo=FALSE}
heights_df <- read.csv("data/r4ds/heights.csv")
head(heights_df)
```

# Fit a linear model
```{r}
earn_lm <-  lm(earn ~ height + sex + ed + age + race, data = heights_df)
earn_lm
```

# View the summary of your model
```{r}
summary(earn_lm)
```

```{r}
predicted_df <- data.frame(
  earn = predict(earn_lm, heights_df),
  ed = heights_df$age, 
  race = heights_df$race, 
  height = heights_df$height,
  age = heights_df$age, 
  sex = heights_df$sex
  )
head(predicted_df)
```

## Compute deviation (i.e. residuals)
```{r}
mean_earn <- mean(heights_df$earn)
mean_earn
```

## Corrected Sum of Squares Total
```{r}
sst <- sum((mean_earn - heights_df$earn)^2)
sst
```

## Corrected Sum of Squares for Model
```{r}
ssm <- sum((mean_earn - predicted_df$earn)^2)
ssm
```

## Residuals
```{r}
residuals <- heights_df$earn - predicted_df$earn
head(residuals)
```

## Sum of Squares for Error
```{r}
sse <- sum(residuals^2)
sse
```

## R Squared
```{r}
r_squared <- ssm / sst
r_squared
```

## Number of observations
```{r}
n <- nrow(heights_df)
n
```

## Number of regression parameters
```{r}
p <- 8
```

## Corrected Degrees of Freedom for Model
```{r}
dfm <- p - 1
dfm
```

## Degrees of Freedom for Error
```{r}
dfe <- n - p
dfe
```

## Corrected Degrees of Freedom Total:   DFT = n - 1
```{r}
dft <- n - 1
dft
```

## Mean of Squares for Model:   MSM = SSM / DFM
```{r}
msm <- ssm / dfm
msm
```

## Mean of Squares for Error:   MSE = SSE / DFE
```{r}
mse <- sse / dfe
mse
```

## Mean of Squares Total:   MST = SST / DFT
```{r}
mst <- sst / dft
mst
```

## F Statistic
```{r}
f_score <- msm / mse
f_score
```

## Adjusted R Squared R2 = 1 - (1 - R2)(n - 1) / (n - p)
```{r}
adjusted_r_squared <- 1 - ((1 - r_squared) * dft) / dfe
adjusted_r_squared
```
# Housing Data

```{r imports, include = FALSE}
library(readxl)
library(dplyr)
library(magrittr)
library(Hmisc)
library(lm.beta)
```

### Data for this assignment is focused on real estate transactions recorded from 1964 to 2016 and can be found in Housing.xlsx. Using your skills in statistical correlation, multiple regression, and R programming, you are interested in the following variables: Sale Price and several other possible predictors.

### If you worked with the Housing dataset in previous week – you are in luck, you likely have already found any issues in the dataset and made the necessary transformations. If not, you will want to take some time looking at the data with all your new skills and identifying if you have any clean up that needs to happen.

```{r housingData, echo = FALSE}
## Load the `data/week-6-housing.csv` to
housing_df = read_excel("data/week-6-housing.xlsx")
housing_df %>% head(4)
```


### i Explain any transformations or modifications you made to the dataset
```{r totalBath, echo = FALSE}
num_baths <- function(bath_full, bath_half, bath_3qtr)
{
    ifelse (bath_full > 0, total <- bath_full * 1, NA)
    
    ifelse (bath_half > 0, total <- total + (bath_half * 0.5), NA)
    
    ifelse (bath_3qtr > 0, total <- total + (bath_3qtr * 0.75), NA)
    
    return (total)
}

housing_df['total_bath_count'] <- num_baths(
    bath_full = housing_df$bath_full_count, 
    bath_half = housing_df$bath_half_count, 
    bath_3qtr = housing_df$bath_3qtr_count)
```

```{r transformations, echo = FALSE}
colnames(housing_df)[2] <- "Sale_Price"
colnames(housing_df)[1] <- "Sale_Date"

housing_df %>% head(4)
```


### ii: Create two variables; one that will contain the variables Sale Price and Square Foot of Lot (same variables used from previous assignment on simple regression) and one that will contain Sale Price and several additional predictors of your choice. Explain the basis for your additional predictor selections.

For my additional predictors, I chose total_bath_count, year_built, square_feet_total_living, and bedroom because they traditionally are used to calculate the price of a house

```{r newVariable1, echo = FALSE}
price_sqft <- select(housing_df, c(Sale_Price, sq_ft_lot))
price_sqft %>% head(5)
```

```{r newVariable2, echo = FALSE}
price_predictors <- select(housing_df, c(Sale_Price, total_bath_count, year_built, 
                                         square_feet_total_living, bedrooms))
price_predictors %>% head(5)
```


### iii: Execute a summary() function on two variables defined in the previous step to compare the model results. What are the R2 and Adjusted R2 statistics? Explain what these results tell you about the overall model. Did the inclusion of the additional predictors help explain any large variations found in Sale Price?
```{r model1, echo = FALSE}
model1 <- lm(Sale_Price ~ sq_ft_lot, data = price_sqft)
summary(model1)
```

The linear regression model for Sales Price and Square Feet per Lot is an 
R^2 = 0.01435 tells us Sq feet per lot only accounts for 1.4% of the variation
in the Sales Price.

```{r model2, echo = FALSE}
model2 <- lm(Sale_Price ~ zip5 + total_bath_count + year_built, + 
               square_feet_total_living + bedrooms, data = price_predictors)
summary(model2)
```

The linear regression model for Sales Price, zip5, total_bath_count, year_built, 
square_feet_total_living, and bedrooms is an R^2 = 0.05 tells us that adding other variables raises the
percentage to 5% in the Sales Price.

The adjusted R^2 with 1 variable is 0.01428 and with 5 variables it is 0.04984 which indicates that adding
the additional variables does help but not a lot.


### iv: Considering the parameters of the multiple regression model you have created. What are the standardized betas for each parameter and what do the values indicate?

```{r, echo = FALSE}
beta_model1 <- lm.beta(model1)
beta_model1
```

Standard Deviation of Sale_Price:

```{r, echo = FALSE}
sd(housing_df$Sale_Price)
```

Standard Deviation of sq_ft_lot:

```{r, echo = FALSE}
sd(housing_df$sq_ft_lot)
```

The standardized beta of 0.12 indicates that as the square footage of the lot increases by 56,933 sq ft,
the Sales Price increases by .12 standard deviations. So for every 56,933 sq ft increase, the Sales price 
goes up by ($404,381 * .12) $48,525.72.

```{r, echo = FALSE}
beta_model2 <- lm.beta(model2)
beta_model2
```
Standard Deviation of zip5:

```{r, echo = FALSE}
sd(housing_df$zip5)
```
The standardized beta of 0.063 indicates that as the zip code changes by 1.70 the Sales increases by 0.063 standard deviations. So for every 1.7 change in zip code, the Sales price goes up by ($404,381 * .63) $254,760.03.

Standard Deviation of total_bath_count:

```{r, echo = FALSE}
sd(housing_df$total_bath_count)
```
The standardized beta of 0.165 indicates that as the bath count increases by 0.70 the Sales increases by 0.165 standard deviations. So for every 0.70 change in bath count, the Sales price goes up by ($404,381 * .165) $66,722.87.

Standard Deviation of year_built:

```{r, echo = FALSE}
sd(housing_df$year_built)
```
The standardized beta of 0.088 indicates that as the year the house was built increases by 17.22, the Sales increases by 0.088 standard deviations. So for every 17.22 increase in the year built, the Sales price goes up by ($404,381 * .088) $35,585.53.

### v: Calculate the confidence intervals for the parameters in your model and explain what the results indicate.
```{r, echo = FALSE}

```

### vi: Assess the improvement of the new model compared to your original model (simple regression model) by testing whether this change is significant by performing an analysis of variance.
```{r, echo = FALSE}

```

### vii: Perform casewise diagnostics to identify outliers and/or influential cases, storing each function's output in a dataframe assigned to a unique variable name.
```{r, echo = FALSE}

```

### viii: Calculate the standardized residuals using the appropriate command, specifying those that are +-2, storing the results of large residuals in a variable you create.
```{r, echo = FALSE}

```

### ix: Use the appropriate function to show the sum of large residuals.
```{r, echo = FALSE}

```

### x: Which specific variables have large residuals (only cases that evaluate as TRUE)?
```{r, echo = FALSE}

```

### xi: Investigate further by calculating the leverage, cooks distance, and covariance rations. Comment on all cases that are problematics.
```{r, echo = FALSE}

```

### xii: Perform the necessary calculations to assess the assumption of independence and state if the condition is met or not.
```{r, echo = FALSE}

```

### xiii: Perform the necessary calculations to assess the assumption of no multicollinearity and state if the condition is met or not.
```{r, echo = FALSE}

```

### xiv: Visually check the assumptions related to the residuals using the plot() and hist() functions. Summarize what each graph is informing you of and if any anomalies are present.
```{r, echo = FALSE}

```

### xv: Overall, is this regression model unbiased? If an unbiased regression model, what does this tell us about the sample vs. the entire population model?
```{r, echo = FALSE}

```

